### 2025-04-17 slides/one_slide/one_slide.tex

\documentclass[10pt]{beamer}
\beamertemplatenavigationsymbolsempty
\usecolortheme{beaver}
\setbeamertemplate{blocks}[rounded=true, shadow=true]
%\setbeamertemplate{footline}[page number]

\usepackage[utf8]{inputenc}
\usepackage[english,russian]{babel}
\usepackage{amssymb,amsfonts,amsmath,mathtext}
\usepackage{subfig}
\usepackage[all]{xy}
\usepackage{array}
\usepackage{multicol}
\usepackage{hyperref}
\usepackage{hhline}
\usepackage{caption}
\usepackage{subfig}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

\graphicspath{ {../figures/} }

%----------------------------------------------------------------------------------------------------------
\title[Loss Landscape Convergence]{Neural Networks Loss Landscape Convergence in Hessian Low-Dimensional Space}
\author[Tem Nikitin \textit{et al.}]{Tem Nikitin \and Nikita Kiselev \and Andrey Grabovoy \and Vladislav Meshkov}
\institute{Moscow Institute of Physics and Technology}
\date{2025}
%----------------------------------------------------------------------------------------------------------
\begin{document}

\begin{frame}{Neural Networks Loss Landscape Convergence in Different Low-Dimensional Spaces}

    \textbf{Goal:} Measure how the loss function changes as the training set size grows:
    $$
        \Delta_k =
        \mathbb{E} \Bigl( \mathcal{L}_{k+1} (\mathbf{w}) - \mathcal{L}_{k} (\mathbf{w}) \Bigr)^2.
    $$

    \textbf{Method:}
    \begin{itemize}
        \item \textbf{Monte Carlo:} Generate points near the minimum according to $p(\mathbf{w})$ and average the differences.
        \item \textbf{Hessian Eigenvectors:} Use directions with the largest eigenvalues to focus on key curvature components.
    \end{itemize}

    \begin{columns}[t]
        \column{0.35\textwidth}
        \centering
        \hspace*{-2cm}
        \includegraphics[width=0.95\textwidth]{img/LS_16.jpg}\\
        \hspace*{-2cm}
        \scriptsize \textit{Loss function}

        \column{0.47\textwidth}
        \centering
        \vspace*{-3.3cm}
        \hspace*{-2.2cm}
        \includegraphics[width=1.6\textwidth]{img/D_32.jpg}\\
        \hspace*{-1.2cm}
        \scriptsize \textit{$\Delta_k$ function}
    \end{columns}

\end{frame}
%----------------------------------------------------------------------------------------------------------
\end{document}