### 2025-04-30 paper/main.tex
Universal commits presents...

\documentclass{article}
\input{new_commands}

% if you need to pass options to natbib, use, e.g.:
\PassOptionsToPackage{numbers, sort, compress}{natbib}
% before loading neurips_2025

% ready for submission
%\usepackage{neurips_2025}

% to compile a preprint version, e.g., for submission to arXiv, add add the [preprint] option:
\usepackage[preprint]{neurips_2025}

% to compile a camera-ready version, add the [final] option, e.g.:
%\usepackage[final]{neurips_2024}

% to avoid loading the natbib package, add option nonatbib:
%\usepackage[nonatbib]{neurips_2024}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors

%%%

\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{xcolor}
\usepackage{textcomp}
\usepackage{manyfoot}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{listings}

\newtheorem{theorem}{Theorem} % continuous numbers
%%\newtheorem{theorem}{Theorem}[section] % sectionwise numbers
%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
\newtheorem{proposition}[theorem]{Proposition}% 
\newtheorem{lemma}{Lemma}% 
%%\newtheorem{proposition}{Proposition} % to get separate numbers for theorem and proposition etc.

\newtheorem{example}{Example}
\newtheorem{remark}{Remark}

\newtheorem{definition}{Definition}
\newtheorem{assumption}{Assumption}

%%%

\title{Neural Networks Loss Landscape Convergence in Hessian Low-Dimensional Space}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{
  Tem Nikitin\\
  Moscow Institute of Physics and Technology\\
  Moscow, Russia\\
  \texttt{nikitin.artem.a@phystech.su}\\
  \And
  Nikita Kiselev\\
  Moscow Institute of Physics and Technology\\
  Moscow, Russia\\
  \texttt{kiselev.ns@phystech.su}\\
  \And
  Andrey Grabovoy\\
  Moscow Institute of Physics and Technology\\
  Moscow, Russia\\
  \texttt{grabovoy.av@phystech.su}\\
  \And
  Vladislav Meshkov\\
  Moscow Institute of Physics and Technology\\
  Moscow, Russia\\
  \texttt{meshkov.ns@phystech.su}\\
}

\begin{document}

\maketitle

\begin{abstract}
  Understanding how a neural network’s loss landscape changes as we add more training data is important for efficient training.
  Although larger datasets are known to reshape this high-dimensional surface, the point when extra data stop making a big difference
  is still unclear.

  In this paper, we study this issue and show that the loss landscape near a local minimum stabilizes once the dataset exceeds a
  certain size. To analyze this, we project the full parameter space onto a smaller subspace formed by the Hessian’s top eigenvectors.
  This low-dimensional view highlights how the loss surface changes in its most important directions. We then apply Monte Carlo sampling
  within this subspace to estimate these changes more precisely.

  We test our approach on standard image‑classification tasks and find that our low-dimensional analysis pinpoints when the landscape
  stops evolving. These findings clarify how dataset size affects optimization and offer practical guidance for balancing training cost
  with performance gains.
\end{abstract}

\textbf{Keywords:}
Neural networks, Loss landscape, Low-dimensional subspace, Hessian eigenvectors, Monte Carlo estimation, Dataset size threshold.

\section{Introduction}\label{sec:intro}

Neural networks are now essential in many areas --- image classification, language modeling, recommender systems, and more. As models and
datasets grow, we often see better accuracy and new breakthroughs. But bigger networks and data also demand more computation time and
resources, which can become prohibitive. While prior work has compared various optimization methods under fixed data and model sizes
\cite{soydaner2020comparison}, the question of when adding more training samples stops yielding significant gains remains largely
unanswered.

In this paper, we investigate exactly that: how the loss landscape changes as we increase the dataset size, and at what point further
data have little effect. Knowing this “minimum viable dataset size” --- a threshold beyond which new samples bring negligible
improvement --- can save both training time and data‑collection effort. We also relate this threshold to generalization behavior, along
the lines of \cite{wu2017towards}.

To study this, we use the Hessian of the loss function as a proxy for local curvature. Computing the full Hessian is expensive, so we
project the parameter space onto a low‑dimensional subspace formed by its top eigenvectors. Our main contributions are:

\begin{enumerate}
  \item Constructing a Hessian‑based projection that captures the principal curvature directions.
  \item Applying Monte Carlo method to identify when the loss landscape stabilizes as dataset size grows.
  \item Visualizing the projected loss surface around local minima to illustrate these changes.
  \item Deriving theoretical criteria --- using results from The Matrix Cookbook \cite{petersen2012matrix} --- to predict when additional
        samples have minimal impact.
\end{enumerate}

This approach is novel in linking low‑rank Hessian estimation to a concrete dataset threshold, enabling more cost‑effective data
collection and a clearer understanding of how data scale interacts with loss geometry. We validate our method on MNIST
\cite{deng2012mnist} and Fashion‑MNIST \cite{xiao2017fashion}, demonstrating how to balance computational cost against accuracy gains.

\textbf{Organization.} The rest of the paper is structured as follows. Section 2 reviews related work. Section 3 introduces notation
and preliminaries. Section 4 derives theoretical bounds on Hessian spectra and the loss‐difference norm. Section 5 presents our
empirical studies. We discuss the implications in Section 6 and conclude in Section 7. Additional experiments and proofs are
provided in Appendix A.

\section{Related Work}\label{sec:rw}

\textbf{Hessian-based Analysis.}
The loss landscape of neural networks has been extensively studied using Hessian-based techniques, which illuminate convergence
behavior and curvature properties \cite{kiselev2024unraveling}. Empirical work shows that near local minima the Hessian often has low
effective rank, indicating a small "active subspace" spanned by a few large eigenvalues \cite{sagun2018empirical}. Ghorbani et al.
\cite{ghorbani2019investigation} examined how these large eigenvalues emerge during training and shape optimization dynamics.
Meshkov et al. \cite{meshkov2024convnets} extended this analysis to fully connected and convolutional architectures, though broader
validation across architectures remains an open challenge.

\textbf{Loss Landscape Geometry and Dataset Size.}
Another research direction explores how dataset size influences model performance and landscape geometry. Networks trained on larger
datasets tend to find flatter, wider minima that generalize better \cite{wu2017towards}. Yet, acquiring and processing massive data is
costly, motivating studies of optimal data–model trade‑offs \cite{hoffmann2022training}. Li et al. \cite{li2018visualizing} developed
visualization tools that reveal how dataset size, initialization, and architecture affect the topology of loss surfaces.

\textbf{Algorithmic Stability and Landscape Sensitivity.}
Algorithmic stability, which measures how small changes in the training set affect the learned model, provides theoretical bounds on
generalization error \cite{bousquet2002stability}. Elisseeff et al. \cite{elisseeff2005stability} extended stability analysis to
randomized algorithms such as bagging and ensemble methods, deriving non-asymptotic generalization guarantees. In our context,
stability theory underpins the intuition that, beyond a certain dataset size, further samples cause negligible changes to the loss
curvature. Our Monte Carlo–based Hessian projection method directly leverages this perspective, using top eigenvectors to quantify
and visualize when the loss landscape stabilizes as new data are added.

\section{Preliminaries}\label{sec:prelim}

\subsection{General notation}

We consider a $S$-class classification problem. Let $\bx \in \bX$ be an input and $y \in \mathcal Y = \{1, \dots, S\}$ its label. A
neural network with parameters $\bw \in \bOmega \subset \mathbb R^M$ defines a mapping $f_{\bw} : \bX \to \mathbb R^S$. Given a dataset
$$
  \mathcal D =
  \{(\bx_i, y_i)\}_{i=1}^N
$$
of $N$ i.i.d. samples, and a per-sample loss $\ell \bigl( f_{\bw}(\bx_i), y_i \bigr)$ (e.g. cross–entropy), we define
$$
  \ell_i(\bw) =
  \ell \bigl( f_{\bw}(\bx_i) , y_i \bigr)\,.
$$

\begin{definition}
  The empirical loss on the first $k$ samples is
  $$
    \cL_k(\bw) =
    \frac1k \sum_{i=1}^k \ell_i(\bw),
  $$
  so that the full-sample loss is $\cL(\bw) = \cL_N(\bw)$.
\end{definition}

A simple telescoping gives the difference
$$
  \cL_{k+1}(\bw) - \cL_k(\bw) =
  \frac{\ell_{k+1}(\bw) - \cL_k(\bw)}{k+1}\,.
$$

\begin{definition}
  The Hessian of $\cL_k$ at $\bw$ is
  $$
    \bH_k(\bw) =
    \nabla^2_{\bw} \cL_k(\bw) =
    \frac1k \sum_{i=1}^k \nabla^2_{\bw} \ell_i(\bw),
  $$
  and, similarly, the full-sample Hessian is $\bH(\bw) = \bH_N(\bw) = \nabla^2_{\bw} \cL_N(\bw)$.
\end{definition}

To capture the overall change in the landscape when adding one more sample, we introduce a $\Delta$‑function:

\begin{definition}
  Let $p(\bw)$ be a weighting (e.g. peaked near a local minimum). Define
  $$
    \Delta_k =
    \int \bigl( \cL_{k+1}(\bw) - \cL_k(\bw) \bigr)^2\, p(\bw)\, d\bw.
  $$
\end{definition}

We study the behavior of $\Delta_k$ as $k \to \infty$ and define a threshold:

\begin{definition}
  Fix a tolerance $\Delta > 0$. The \emph{$\Delta$-sufficient sample size} is
  $$
    K =
    \inf \{ k \mid \forall\, k' \ge k : \; \Delta_{k'} < \Delta\}.
  $$
  When $k \ge K$, adding further samples changes the loss landscape by less than $\Delta$.
\end{definition}

\subsection{Assumptions}

We make the following standard assumptions to simplify our analysis around a local minimizer $\bw^*$.

\begin{assumption}
  The point $\bw^*$ is a common local minimizer for both $\cL_k$ and $\cL_{k+1}$, i.e.
  $$
    \nabla \cL_k(\bw^*) =
    \nabla \cL_{k+1}(\bw^*) =
    0.
  $$
\end{assumption}

Under this assumption, a second‑order Taylor expansion of $\cL_k$ about $\bw^*$ gives
$$
  \cL_k(\bw) \approx
  \cL_k(\bw^*) + \tfrac12\, (\bw - \bw^*)^{\T} \bH_k(\bw^*)\, (\bw - \bw^*).
$$

\begin{assumption}
  At the minimizer $\bw^*$, the loss values for consecutive sample sizes are the same:
  $$
    \cL_{k+1}(\bw^*) =
    \cL_k(\bw^*).
  $$
\end{assumption}

\begin{assumption}
  The parameters $\bw$ are distributed according to a density $p(\bw)$ (e.g. a Gaussian prior centered at $\bw^*$). Then
  $$
    \Delta_k =
    \int \bigl(\cL_{k+1}(\bw)-\cL_k(\bw)\bigr)^2\, p(\bw)\, d\bw =
    \mathbb{E} \left( \cL_{k+1}(\bw) - \cL_{k}(\bw) \right)^2 =
  $$
  $$
    \mathbb D \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr] +
    \Bigl( \mathbb E \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr] \Bigr)^2.
  $$
\end{assumption}

\section{Method: Projection onto Dominant Eigen-Directions and Loss Landscape Approximation}\label{sec:meth}

Working with the full Hessian $\mathbf{H}_k(\bw)$ of a neural network is computationally expensive due to the high dimensionality
of the parameter space $\bOmega$. To reduce this cost, we project the parameters onto a lower‑dimensional subspace spanned by the
$D$ dominant eigenvectors of the Hessian at a local minimum $\bw^*$.

The justification for this projection is that, empirically, the Hessian has low effective rank near a minima \cite{sagun2018empirical}:
only a few eigenvalues are significantly above zero, while the rest are negligible. Therefore, directions corresponding to small
eigenvalues contribute little to loss variation. Retaining only the top $D$ eigenvectors yields a compact yet informative representation
of the most important curvature directions.

For illustration, Figure \ref{fig:evgen} plots the top few eigenvalues of $\bH(\bw^*)$ in descending order. We compute these dominant
eigenvalues using the iterative power‐method described at \cite{hessian-eigenthings}.

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.7\textwidth]{img/eigenvalues.pdf}
  \caption{Top eigenvalues of $\bH(\bw^*)$ in descending order.}
  \label{fig:evgen}
\end{figure}

Concretely, let
$$
  \bP =
  \bigl[ \mathbf{e}_1, \dots, \mathbf{e}_D \bigr]
$$
be the matrix whose columns are the top $D$ eigenvectors. We write any parameter vector as
$$
  \bw =
  \bw^* + \bP\, \btheta, \quad
  \btheta \in \mathbb R^D.
$$
Substituting into a second‑order Taylor expansion of $\cL_k$ around $\bw^*$ gives
$$
  \cL_k\bigl(\bw^* + \bP\,\btheta\bigr) \approx
  \cL_k(\bw^*) + \tfrac12\, \btheta^{\T}\! \bigl( \bP^{\T}\, \bH_k(\bw^*)\, \bP\bigr)\, \btheta =
  \cL_k(\bw^*) + \tfrac12\, \btheta^{\T} \bLambda_k\, \btheta,
$$
where
$$
  \bLambda_k =
  \bP^{\T}\, \bH_k(\bw^*)\, \bP =
  \operatorname{diag} \bigl( \lambda_k^{(1)}, \dots, \lambda_k^{(D)} \bigr).
$$
An identical expansion holds for $\cL_{k+1}$, yielding $\bLambda_{k+1}$ from the Hessian on $k+1$ samples.

\begin{theorem}[Approximate $\Delta_k$ via Eigenvalues]\label{th:th_1}
  Under the assumption that $\btheta \sim \mathcal N(0, \sigma^2 \mathbf I_D)$, the change in loss can be estimated by
  $$
    \Delta_k =
    \mathbb E \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr]^2 \approx
    \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{d=1}^D (\lambda_{k+1}^{(d)} - \lambda_k^{(d)})^2
    + \Bigl( \sum_{d=1}^D (\lambda_{k+1}^{(d)} - \lambda_k^{(d)}) \Bigr)^2 \Biggr).
  $$
\end{theorem}

\begin{remark}
  Assuming $\mathbb{E}(\btheta) = \mathbf{0}$ is natural when centering the sampling distribution at the minimum.
\end{remark}

This formula uses only the top $D$ eigenvalues, so it is efficient to compute even for very large models.
In practice, it provides an upper bound on the empirical $\Delta_k$ and thus a reliable criterion for determining when the
dataset size is sufficient.

\section{Experiments}\label{sec:exp}

To validate our theoretical estimates, we perform a set of experiments that track how the loss landscape changes as we grow the
training set.

\paragraph{Datasets.}
We evaluate on MNIST \cite{deng2012mnist} and Fashion‑MNIST \cite{xiao2017fashion}, each containing 60,000 training and 10,000 test
grayscale images of size $28\times28$ pixels.

\paragraph{Model Architecture.}
All experiments use a simple multilayer perceptron (MLP) with ReLU activations. We keep the same architecture—number of hidden layers
and hidden units—across all runs to ensure a fair comparison.

\paragraph{Experimental Procedure.}
\begin{enumerate}
  \item \textbf{Random‐direction visualization \ref{sec:exp:pre}:} As an initial sanity check, we project the parameters onto two random directions and
        plot the resulting 3D loss landscapes, their squared differences, and the corresponding $\Delta_k$ for various settings.
  \item \textbf{Hessian‐based analysis \ref{sec:exp:hes}:} We repeat the above visualization using the top eigenvectors of the Hessian. This directly
        tests our hypothesis regarding the dominant role of the largest Hessian eigenvalues.
  \item \textbf{Main experiment \ref{sec:exp:main}:} We compute both the theoretical and empirical values of $\Delta_k$ and compare them. This evaluates
        the accuracy of our approximation predicted by theory.
\end{enumerate}

All code and detailed configuration files are available in our GitHub
repository \footnote{\href{https://github.com/intsystems/2025-Project-183}{GitHub}}.
\subsection{Preliminary Experiment: Projection onto Random Directions}\label{sec:exp:pre}
\subsubsection{Visualizing the Loss Landscape}

To build intuition for how the loss surface changes with training‐set size, we first project the model parameters onto two random
directions. Specifically, we pick two random vectors $\mathbf v_1, \mathbf v_2 \in \mathbb R^P$ in parameter space and study the
subspace they span.

We then evaluate the empirical loss
$$
  \cL_k(\bw) =
  \frac1k \sum_{i=1}^k \ell_i(\bw),
$$
where $\ell_i(\bw) = \ell (f_\bw(\bx_i), y_i)$, over a grid in the $(\alpha, \beta)$ plane defined by
$$
  \bw(\alpha,\beta) =
  \bw^* + \alpha\, \mathbf v_1 + \beta\, \mathbf v_2, \quad
  (\alpha, \beta) \in [-1,1]^2.
$$

First, a uniform grid of $(\alpha,\beta)$ pairs on the domain $[-1,1]\times[-1,1]$ with fixed step size $h$ is constructed.
For each grid point $(\alpha,\beta)$, we set $\mathbf{w} = \mathbf{w}(\alpha,\beta)$, keeping the base weights
$\mathbf{w}^*$ fixed, and compute the losses $\mathcal{L}_k(\mathbf{w})$ and $\mathcal{L}_{k+1}(\mathbf{w})$.
This produces two scalar fields over the grid: the first, $\mathcal{L}_k(\alpha,\beta)$, illustrates how the mean loss
varies across the parameter space, while the second,
$$\bigl(\mathcal{L}_{k+1}(\alpha,\beta) - \mathcal{L}_k(\alpha,\beta)\bigr)^2,$$ highlights the regions where adding a single
sample induces the largest change. Finally, both fields are visualized as two adjacent 3D plots.

\begin{figure}[!htbp]
  \hspace*{-2.2cm}
  \subfloat{\includegraphics[width=1.3\textwidth]{img/loss_random_1_2.pdf}}
  \caption{$\cL_k$ (left) and $\bigl[ \cL_{k+1}(\btheta) - \cL_k(\btheta) \bigr]^2$ (right) for small $k$, projected onto the two random vectors.}
  \label{fig:loss_random_small}
\end{figure}

\begin{figure}[!htbp]
  \hspace*{-2.2cm}
  \subfloat{\includegraphics[width=1.3\textwidth]{img/loss_random_-2_-1.pdf}}
  \caption{$\cL_k$ (left) and $\bigl[ \cL_{k+1}(\btheta) - \cL_k(\btheta) \bigr]^2$ (right) for maximum $k$, projected onto the two random vectors.}
  \label{fig:loss_random_big}
\end{figure}

\paragraph{Loss Landscape for the Small Dataset}
In Figure \ref{fig:loss_random_small}, the loss surface is highly irregular, with sharp peaks and deep valleys.
This irregularity indicates that with very few training samples, the model is extremely sensitive to small parameter perturbations.
The squared difference surface also exhibits large spikes, confirming that adding a single data point can cause substantial local changes
in the loss.

\paragraph{Loss Landscape for the Full Dataset}
In contrast, Figure \ref{fig:loss_random_big} shows a much smoother loss surface with gentle slopes. The overall loss
values are lower --- reflecting the averaging effect of more samples --- and the squared difference surface is correspondingly much smaller.
This smoothing demonstrates that as the dataset grows, the local optimum stabilizes, and additional samples produce only minor
adjustments to the loss landscape. Hence, beyond a certain dataset size, further data have negligible impact on the shape of the loss
surface.

\subsubsection{Visualizing the \texorpdfstring{$\Delta_k$}{Delta k} Function}

To study how the loss landscape converges as the dataset grows, we estimate and plot the function
$$
  \Delta_k =
  \mathbb{E} \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr]^2
$$
via a Monte Carlo approximation using Gaussian perturbations in parameter space.

Starting from a fixed trained model to get $\mathbf{w}^*$, we draw $T$ independent random direction vectors
$$
  \btheta^{(t)} \sim \mathcal{N}(\mathbf{0}, \sigma^2 \mathbf{I}_d), \quad t = 1, \dots, T.
$$
For each batch, we form a perturbed parameter set $\mathbf{w}^{(t)} = \mathbf{w}^* + \btheta^{(t)}$,
and evaluate the losses $\mathcal{L}_k(\mathbf{w}^{(t)})$ for each $k = 1, \, \dots, \, N$,
after what compute the squared difference on given values:
$$
  d_t = \bigl[\mathcal{L}_{k+1}(\mathbf{w}^{(t)}) - \mathcal{L}_k(\mathbf{w}^{(t)})\bigr]^2.
$$
We then approximate the average change
$\Delta_k \approx \frac{1}{T} \sum_{t=1}^T d_t$,
noting that increasing $T$ reduces the variance of this Monte Carlo estimator. Finally, we plot $\Delta_k$ against $k$ to observe its decay
and include the rescaled quantity $\Delta_k \cdot k^2$ to highlight the expected $\mathcal{O}(1/k)$ convergence rate.

Figures \ref{fig:delta_random_sigma} and \ref{fig:delta_random_dim} present the behavior of the function
$$
  \Delta_k =
  \mathbb{E} \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr]^2
$$
under different Monte Carlo sampling settings and projection subspace sizes. Each figure isolates one variable:

\begin{itemize}
  \item Figure \ref{fig:delta_random_sigma}: varying the Gaussian variance $\sigma$.
  \item Figure \ref{fig:delta_random_dim}: varying the subspace dimension $D$.
\end{itemize}

\begin{figure}[!htbp]
  \hspace*{-2.6cm}
  \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_random_sigma_2_1024.pdf}}
  \caption{$\Delta_k$ (left) and $\Delta_k \cdot k^2$ (right), varying the Gaussian variance with fixed $d = 2$.}
  \label{fig:delta_random_sigma}
\end{figure}

\begin{figure}[!htbp]
  \hspace*{-2.6cm}
  \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_random_dim_1_1024.pdf}}
  \caption{$\Delta_k$ (left) and $\Delta_k \cdot k^2$ (right), varying the subspace dimension with fixed $\sigma = 1$.}
  \label{fig:delta_random_dim}
\end{figure}

\paragraph{Comparing Different Variances \texorpdfstring{$\sigma$}{sigma}}
Figure \ref{fig:delta_random_sigma} shows that a high Gaussian variance $\sigma$ produces a significant change in both the shape
and magnitude of the $\Delta_k$ curves. This occurs because, as we move away from the local minimum, the value of the loss
function changes. However, this change is modest: the insensitivity likely stems from using random projection directions,
which do not align with the dominant curvature axes of the Hessian, so scaling those directions has little impact on the estimated
loss differences. Moreover, variances not exceeding one yield very similar effects, so we choose $\sigma = 1$ for simplicity in our
calculations.

\paragraph{Comparing Different Subspace Dimensions}
Figure \ref{fig:delta_random_dim} shows that, when using random directions, increasing the projection dimension $D$ does not
significantly alter the overall shape of $\Delta_k$. This suggests that only a few directions dominate the variation --- motivating our
focus on Hessian‐based projections in the main experiment.

In all cases, the right‐hand plots (showing $\Delta_k \cdot k^2$) exhibit a roughly $\mathcal O(1 / k^2)$ decay, confirming that the
incremental effect of adding one more sample diminishes proportionally to $1 / k$ as the dataset grows.

The preliminary experiments show that the loss landscape indeed smooths out and becomes more stable as the training set size $N$
increases. However, because we used random projection directions, these results capture only a generic view of stability. This
observation motivates our main experiment, in which we project onto the Hessian’s top eigenvectors --- i.e., the directions of greatest
curvature --- to obtain a more precise understanding of when and how the loss surface converges.

\subsection{Moving Forward: Projection onto Dominant Hessian Eigenvectors}\label{sec:exp:hes}

\paragraph{Projection Directions.}
In the main experiment, we replace random directions with the $D$ eigenvectors of the Hessian at the local minimizer $\bw^*$ that
correspond to its largest eigenvalues. Prior work \cite{sagun2018empirical} shows these top modes capture nearly all of the
non‑negligible curvature, and Figure \ref{fig:evgen} confirms a rapid eigenvalue decay. Fixing $D = 10$ balances computational cost and
fidelity, and our tests on Fashion‑MNIST demonstrate that this choice suffices even for complex architectures.

\paragraph{Objectives.}
By projecting onto these principal directions, we aim to:
\begin{itemize}
  \item Obtain a clearer view of how the loss surface changes as $k$ grows, since these directions drive the largest loss variations.
  \item Accurately identify the threshold $k^*$ at which further data cause only negligible shifts in the local curvature.
  \item Empirically validate our theoretical $\Delta_k$ estimates by comparing them against Monte Carlo measurements in this eigenvector
        subspace.
\end{itemize}

\subsubsection{Visualizing the Loss Landscape}

Figures \ref{fig:loss_eigen_small} and \ref{fig:loss_eigen_big} display 3D plots of $\cL_k(\btheta)$ (left) and
$\bigl[ \cL_{k+1}(\btheta) - \cL_k(\btheta) \bigr]^2$ (right), where $\btheta$ varies over the span of the top two Hessian eigenvectors.
The procedure mirrors the random‑direction case but uses these maximally informative axes.

\begin{figure}[!htbp]
  \hspace*{-2.2cm}
  \subfloat{\includegraphics[width=1.3\textwidth]{img/loss_eigen_1_2.pdf}}
  \caption{$\cL_k$ (left) and $\bigl[ \cL_{k+1}(\btheta) - \cL_k(\btheta) \bigr]^2$ (right) for small $k$, projected onto the top two Hessian eigenvectors.}
  \label{fig:loss_eigen_small}
\end{figure}

\begin{figure}[!htbp]
  \hspace*{-2.2cm}
  \subfloat{\includegraphics[width=1.3\textwidth]{img/loss_eigen_-2_-1.pdf}}
  \caption{$\cL_k$ (left) and $\bigl[ \cL_{k+1}(\btheta) - \cL_k(\btheta) \bigr]^2$ (right) for maximum $k$, projected onto the top two Hessian eigenvectors.}
  \label{fig:loss_eigen_big}
\end{figure}

\paragraph{Loss Landscape for the Small Dataset}
In Figure \ref{fig:loss_eigen_small}, the loss surface projected onto the top two Hessian eigenvectors is
noticeably more structured and less erratic than in the random‐direction case. This confirms that these eigenvectors capture the
most influential directions of loss variation. The squared‐difference plot also shows larger peaks, indicating that the model is
highly sensitive to new data along these key axes when $k$ is small.

\paragraph{Loss Landscape for the Full Dataset}
By contrast, Figure \ref{fig:loss_eigen_big} shows a much smoother loss surface and a greatly reduced squared‐difference.
As the dataset grows, the loss changes less along the principal curvature directions, and the model’s local behavior stabilizes.
Notably, even at large $k$, the squared difference remains higher than in the random‐direction experiment, providing a practical upper
bound on landscape fluctuations. These findings reinforce that projecting onto the top eigenvectors effectively reveals the convergence
behavior of the loss landscape and helps intuitively understand the existence of the point beyond which additional data have minimal impact.

\subsubsection{Visualizing the \texorpdfstring{$\Delta_k$}{Delta k} Function}

\begin{figure}[!htbp]
  \hspace*{-2.6cm}
  \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_eigen_1_10_1024.pdf}}
  \caption{$\Delta_k$ (left) and $\Delta_k \cdot k^2$ (right)}
  \label{fig:delta_eigen}
\end{figure}

Figure \ref{fig:delta_eigen} presents the Monte Carlo estimate of $\Delta_k$ in the subspace spanned by the top $d = 10$
Hessian eigenvectors. Compared to the random‐direction experiment, $\Delta_k$ here has a higher value --- reflecting the
large curvature captured by these eigenvectors. The right‐hand plot of $\Delta_k\cdot k^2$
confirms an approximate $\mathcal O(1 / k^2)$ convergence rate. These observations show that projecting onto the leading eigen-directions not only
highlights the greatest loss sensitivity but also yields a reliable upper bound on how quickly the landscape stabilizes as more data are
added.

\subsection{Main Experiment: Theoretical vs. Empirical \texorpdfstring{$\Delta_k$}{Delta k}}\label{sec:exp:main}

In this experiment, we compare our Hessian‐based theoretical estimate of $\Delta_k$ with the empirical Monte Carlo measurements.

As the Theorem \ref{th:th_1} says, we approximate $\Delta_k$ by
$$
  \Delta_k \approx
  \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{d=1}^D (\lambda_{k+1}^{(d)} - \lambda_k^{(d)})^2
  + \Bigl( \sum_{i=1}^D (\lambda_{k+1}^{(d)} - \lambda_k^{(d)}) \Bigr)^2 \Biggr).
$$
where $\{\lambda_k^{(i)}\}_{i=1}^d$ and $\{\lambda_{k+1}^{(i)}\}_{i=1}^d$ are the top $D$ Hessian eigenvalues at sample sizes $k$ and $k + 1$,
and $\sigma$ is the perturbation variance.

We compute both empirical and theoretical estimates of $\Delta_k$. First, we train a model instance as described above.
For each batch $k$, we apply the power‐iteration method to the model to extract the top $D$ Hessian eigenvalues
$\{\lambda_k^{(d)}\}_{d=1}^D$. The theoretical estimate of $\Delta_k$ is then obtained by substituting these eigenvalues into our
approximation. To reduce high-frequency fluctuations, we apply a five‐point moving average to the sequence:
$\frac{1}{5}\bigl(\Delta_{k-2} + \Delta_{k-1} + \Delta_k + \Delta_{k+1} + \Delta_{k+2}\bigr)$ for each $k$,
where, at the boundaries, any index outside $[1,N]$ is replaced by the nearest valid endpoint (i.e.\ we replicate the border value),
which is equivalent to convolution with constant‐value padding.

In parallel, we compute empirical estimates of $\Delta_k$ via Monte Carlo sampling of Gaussian perturbations along the same
top-$D$ eigenvectors, following the procedure described above. Finally, we plot the
theoretical and empirical $\Delta_k$ on the same axes, allowing us to assess how closely the eigenvalue-based bound tracks the observed
convergence of the loss landscape.

\begin{figure}[!htbp]
  \subfloat{\includegraphics[width=1\textwidth]{img/delta_border_1_10_1024.pdf}}
  \caption{Empirical and Theoretical $\Delta_k$}
  \label{fig:delta_eigen_comparison}
\end{figure}

Figure \ref{fig:delta_eigen_comparison} shows the theoretical estimate of $\Delta_k$ (dashed line) alongside the empirical Monte Carlo
measurements (solid line). Across all values of $k$, the theory curve remains above the empirical one, demonstrating that our
eigenvalue‐based formula provides a valid upper bound for the true loss‐landscape changes.

Now, to eliminate the explicit batch index $k$, we plot $\Delta_k$ directly against $\cL_k$ and apply the same moving average
for a smoother, more stable curve.

Figure \ref{fig:delta_loss} shows the relationship between the loss $\cL_k(\bw^*)$ and the corresponding change $\Delta_k(\bw^*)$.
In the top panel, each blue dot represents the empirical estimate of $\Delta_k$ versus $\cL_k$, and the red line is the best-fit
linear trend. The bottom panel shows the analogous theoretical estimates, again smoothed and fitted (green line). In both cases,
there is a strong positive correlation: higher loss values correspond to larger observed and predicted changes. Although some
scatter remains due to sampling noise, the overall linear relationship validates our hypothesis that the eigenvalue-based
bound $\Delta$ captures the dominant curvature effects of the Hessian. Consequently, the $\Delta$-sufficient condition can effectively
limit changes in the loss function.

\begin{figure}[!htbp]
  \subfloat{\includegraphics[width=1\textwidth]{img/delta_loss_1_10_1024.pdf}}
  \caption{Empirical and Theoretical $\Delta_k$}
  \label{fig:delta_loss}
\end{figure}

We are now ready to propose an algorithm to determine a $\Delta$-sufficient dataset size. Both theoretically and empirically, once
$\Delta_k$ falls below a prescribed tolerance $\Delta$, the loss $\cL_k$ changes negligibly. Thus we say that
for a tolerance $\Delta > 0$. The \emph{$\Delta$-sufficient sample size} is
$$
  K =
  \inf \{ k \mid \forall\, k' \ge k : \; \Delta_{k'} < \Delta\}.
$$
When $k \ge K$, adding further samples changes the loss landscape by less than $\Delta$.
Smaller values of $\Delta$ correspond to higher accuracy. In our MNIST experiments, we found that
$\Delta=0.025$ yields $k=100$ and $\cL_k\approx0.013$.

Table~\ref{tab:results} summarizes the key results of our $\Delta$-sufficient criterion applied to MNIST and Fashion-MNIST.
For each dataset and model architecture, we report the tolerance $\Delta$, the number of batches $k$ required to reach
$\Delta$-sufficiency, the resulting loss $\mathcal{L}_k$, and the total runtime.

\begin{table}[ht]
  \centering
  \caption{Results of $\Delta$-sufficient experiments on MNIST and Fashion-MNIST}\label{tab:results}
  \begin{tabular}{llcccc}
    \toprule
    Dataset       & Model             & $\Delta$ & $k$ & $\mathcal{L}_k$ & Time (s) \\
    \midrule
    MNIST         & Single-layer MLP  & 0.025    & 100 & 0.013           & 2000     \\
                  & Multi-layer MLP   & 0.025    & 40  & 0.010           & 3500     \\
                  & Convolutional CNN & 0.025    & 60  & 0.024           & 1800     \\
    \midrule
    Fashion-MNIST & Single-layer MLP  & 0.030    & 120 & 0.020           & 2100     \\
                  & Multi-layer MLP   & 0.030    & 90  & 0.017           & 4400     \\
                  & Convolutional CNN & 0.030    & 70  & 0.015           & 2400     \\
    \bottomrule
  \end{tabular}
\end{table}

Since our theoretical estimate $\Delta_k$ consistently bounds the
empirical $\Delta_k$ and is much cheaper to compute, we employ the following algorithm for stopping criterion:

\begin{algorithm}[H]
  \caption{Determine $\Delta$-sufficient dataset size}\label{alg:algo}
  \begin{algorithmic}[1]
    \Require target tolerance $\Delta>0$, batch size $B$
    \State $\mathcal{D}\gets \emptyset,\;k\gets 0$
    \Repeat
    \State $k\gets k+1$
    \State Sample a new batch of $B$ examples and append to $\mathcal{D}$
    \State Train or update the model on $\mathcal{D}$
    \State Compute the top-$D$ Hessian eigenvalues and form
    $\Delta_k$ estimation via the eigenvalue-based approximation:
    $$
      \Delta_k \approx
      \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{d=1}^D (\lambda_{k+1}^{(d)} - \lambda_k^{(d)})^2
      + \Bigl( \sum_{i=1}^D (\lambda_{k+1}^{(d)} - \lambda_k^{(d)}) \Bigr)^2 \Biggr).
    $$
    \Until{$\Delta_k > \Delta$}
    \State $\textbf{return }|\mathcal{D}|$
  \end{algorithmic}
\end{algorithm}

Moreover, in addition to determining the dataset size $K$, the procedure also yields a fully trained model as a by‐product of the
iterative updates.

This procedure incrementally grows the dataset until the eigenvalue-based estimate $\Delta_k$ falls below the prescribed
tolerance $\Delta$, at which point further additions have a negligible effect on the loss. The dominant cost per iteration is the
extraction of the top-$D$ Hessian eigenvalues via power iteration, which requires $\mathcal O(I\,D\,C_{\mathrm{Hv}})$ work where $I$
is the number of power‐iteration steps and $C_{\mathrm{Hv}}$ is the cost of a single Hessian–vector product (comparable to one
gradient evaluation). Since $D\ll M$ and $I$ is typically small (e.g.\ $I\approx 20$), the extra overhead
remains linear in the number of mini-batches and is modest compared to full second-order methods. Thus, our algorithm offers an
efficient, theoretically grounded approach to determining a $\Delta$-sufficient dataset size with minimal computational burden.

\section{Discussion}\label{sec:disc}

The gap between the theoretical and empirical curves arises because the estimate uses only the top $D$ eigenvalues. Therefore, focusing on
the leading modes yields a reliable and efficient criterion for determining when adding more data has negligible effect on the local loss
geometry.

Our Hessian‐projection framework provides a clear picture of how a neural network’s loss landscape evolves with training–set size.
By restricting attention to the $D$ principal curvature directions (the top Hessian eigenvectors), we obtain far more structured
3D loss surfaces and a sharply decaying landscape‐difference measure
$$
  \Delta_k =
  \mathbb{E}\bigl[\cL_{k+1}(\btheta) - \cL_k(\btheta)\bigr]^2,
$$
which empirically follows an approximate $\mathcal{O}(1/k^2)$ trend. Moreover, the Hessian‐based theoretical bound
consistently upper‐bounds the Monte Carlo estimates (see Figure \ref{fig:delta_eigen_comparison}). This confirms that only a small number
of eigen‐modes drive the bulk of loss‐landscape change.

A practical outcome is a stopping criterion and algorithm \ref{alg:algo} for data collection (see Figure \ref{fig:delta_loss}):
once $\Delta_k$ falls below a fixed tolerance $\Delta$, further additions have a negligible effect on the loss. This lets practitioners
avoid unnecessary data collection and computation without sacrificing model stability.

\paragraph{Limitations and Future Work}
Our study used $D = 10$ principal directions and relatively small MLPs on MNIST‐like tasks. Deeper or convolutional architectures,
alternative loss functions, or non–i.i.d. data may shift the spectrum decay and require reconsidering $D$. Extending this
Hessian‐based convergence criterion to large models or streaming data remains an exciting direction.

\section{Conclusion}\label{sec:concl}

We have introduced and validated a Hessian‐projection approach to identify when a neural network’s loss landscape “settles” as
training data grows. By projecting onto the top eigenvectors of the Hessian, we both visualize and quantify the landscape’s
sensitivity via the $\Delta_k$ function, and derive a tight eigenvalue‐difference bound that serves as a reliable upper bound on
loss‐landscape changes. Our results demonstrate a clear “sufficient sample size” threshold: beyond this, further data become
insignificant in local curvature. This criterion provides a concrete guide for efficient dataset sizing and early stopping in
model training.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\nocite{*}

\bibliographystyle{unsrtnat}
\bibliography{references}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\appendix
\section{Appendix}\label{app}

\subsection{Proof of Theorem 1}\label{app:th_1}

\begin{proof}
  Starting from the second‐order Taylor expansion around the local minimizer $\bw^*$, we have
  $$
    \cL_{k+1}(\bw) - \cL_k(\bw) \approx
    \bigl[ \cL_{k+1}(\bw^*) - \cL_k(\bw^*) \bigr] + \tfrac12\, \btheta^{\T}\! \bigl( \bLambda_{k+1} - \bLambda_k \bigr)\, \btheta,
  $$
  where $\btheta = \bP^{\T} (\bw-\bw^*)$ and $\bLambda_j = \bP^{\T} \bH_j(\bw^*) \bP$.

  Let $p(\btheta)$ be a distribution with mean $\bm=\mathbb E[\btheta] = 0$ and covariance $\bSigma = \mathbb D[\btheta]$. Denote
  $$
    \bA=\bLambda_{k+1} - \bLambda_k.
  $$
  Then, by standard matrix‐expectation results (see \cite[p. 35]{petersen2012matrix}),
  $$
    \mathbb E \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr] =
    \mathbb E \bigl[ \tfrac12\, \btheta^{\T} \bA\, \btheta \bigr] =
    \tfrac12\, \bigl( \tr(\bA \bSigma) + \bm^{\T} \bA \bm \bigr) =
    \tfrac12\, \tr(\bA \bSigma).
  $$
  Similarly, using \cite[p. 43]{petersen2012matrix},
  $$
    \mathbb D \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr] =
    \tfrac14 \Bigl( 2\, \tr(\bA^2 \bSigma^2) + 4\, \bm^{\T} \bA^2 \bm\, \sigma^2 \Bigr) =
    \tfrac12\, \sigma^4\, \tr(\bA^2),
  $$
  where we have set $\bSigma=\sigma^2\mathbf I$ and used $\bm=0$.

  Since
  $$
    \Delta_k =
    \mathbb D \bigl[ \cL_{k+1} - \cL_k \bigr] + \bigl( \mathbb E [\cL_{k+1} - \cL_k] \bigr)^2,
  $$
  we get
  $$
    \Delta_k \approx
    \tfrac12\, \sigma^4\, \tr(\bA^2) + \tfrac14\, \bigl( \tr(\bA \bSigma) \bigr)^2 =
    \tfrac{\sigma^4}{4} \Bigl( 2\, \tr(\bA^2) + \bigl( \tr(\bA) \bigr)^2 \Bigr).
  $$
  Finally, since
  $$
    \tr(\bA) =
    \sum_{d=1}^D (\lambda_{k+1}^d - \lambda_k^d), \quad
    \tr(\bA^2) =
    \sum_{d=1}^D (\lambda_{k+1}^d - \lambda_k^d)^2,
  $$
  we arrive at the stated approximation
  $$
    \Delta_k \approx
    \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{d=1}^D (\lambda_{k+1}^{(d)} - \lambda_k^{(d)})^2
    + \Bigl( \sum_{d=1}^D (\lambda_{k+1}^{(d)} - \lambda_k^{(d)}) \Bigr)^2 \Biggr).
  $$
\end{proof}

\end{document}