### 2025-05-22 slides/inn_prac/thesis/dd.tex

\documentclass[a4paper,12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\geometry{margin=2.5cm}

\begin{document}

\noindent\textbf{UDC 519.688.4}\\[2ex]
\begin{center}
  {\LARGE \textbf{Neural Networks Loss Landscape Convergence in Hessian Low-Dimensional Space}}\\[2ex]
  Tem Nikitin$^{1}$, Nikita Kiselev$^{1}$, Andrey Grabovoy$^{1}$, Vladislav Meshkov$^{1}$\\[1ex]
  $^{1}$Moscow Institute of Physics and Technology, Moscow, Russia
\end{center}

\section*{Thesis}

Deep neural networks often improve when we feed them more data, because extra samples make the loss surface smoother and help the model generalize. But after a certain point, new data barely changes the local shape of the loss, and collecting more can waste time and resources. We need a way to find the minimal amount of data that still shapes the model’s behaviour well.

We offer a simple bound (Theorem~1) for the average squared change in loss:
$$
\Delta_k = \mathbb{E}[L_{k+1}(w) - L_k(w)]^2 \approx \frac{\sigma^4}{4} \Bigl(2\sum_{d=1}^D(\lambda_{k+1}^{(d)} - \lambda_k^{(d)})^2 + \bigl(\sum_{d=1}^D(\lambda_{k+1}^{(d)} - \lambda_k^{(d)})\bigr)^2\Bigr),
$$
where $\lambda_k^{(d)}$ are the top $D$ eigenvalues of the Hessian at batch size $k$, and $\sigma^2$ measures random perturbations. This formula works in a low-dimensional space spanned by the main curvature directions.

Our method runs as follows:
\begin{itemize}
  \item Use power iterations to find the top $D=10$ Hessian eigenvectors.
  \item After each new mini-batch, project updates onto this subspace and compute both the actual $\Delta_k$ by sampling and the bound above.
  \item Stop adding data once the bound drops below a chosen tolerance $\Delta$.
\end{itemize}

We tested on MNIST and Fashion-MNIST with simple MLP and CNN models. Figure~\ref{fig:delta} shows that the theoretical bound stays above the measured $\Delta_k$ curves and shrinks roughly like $O(1/k)$. Table~1 lists the minimal batch sizes $K$ needed for tolerances of 0.025 and 0.030. Values range from 40 to 120, and more complex models converge faster because they capture sharper directions.

Each step costs $O(I\,D\,C_{Hv})$ operations, where $I$ is power-iteration steps and $C_{Hv}$ is the cost of a Hessian-vector product. We also discuss simple tricks, like reducing $D$ or stopping iterations early, to speed this up.

In future work, we plan to apply this to deeper, non-convex networks, data streams, and active sampling. Embedding this check in online learning could let systems decide in real time whether to ask for more data.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.8\textwidth]{figure9.png}
  \caption{Empirical and theoretical $\Delta_k$ for MNIST experiments.}
  \label{fig:delta}
\end{figure}

\begin{table}[h]
  \centering
  \caption{Δ-sufficient sample size results}
  \begin{tabular}{lccc}
    \hline
    Dataset       & Model & $\Delta$ & $K$ \\
    \hline
    MNIST         & MLP   & 0.025    & 100 \\
    MNIST         & CNN   & 0.025    & 60  \\
    Fashion-MNIST & CNN   & 0.030    & 70  \\
    \hline
  \end{tabular}
\end{table}

\paragraph{Funding and Acknowledgements}
This work was supported by the Moscow Institute of Physics and Technology. We thank the HPC Center for computing resources and helpful discussions on Hessian analysis.

\begin{thebibliography}{9}
  \bibitem{soy2020} D. Soydaner, \textit{A comparison of optimization algorithms for deep learning}, Int. J. Pattern Recognit. Artif. Intell., 34(13):2052013, 2020.
  \bibitem{wu2017} L. Wu, Z. Zhu, W. E., \textit{Towards understanding generalization of deep learning: Perspective of loss landscapes}, arXiv:1706.10239, 2017.
  \bibitem{cookbook} K. B. Petersen, M. S. Pedersen, \textit{The Matrix Cookbook}, 2012.
  \bibitem{deng2012} L. Deng, \textit{The MNIST database of handwritten digit images for machine learning research}, IEEE Signal Process. Mag., 29(6), 2012.
  \bibitem{xiao2017} H. Xiao, K. Rasul, R. Vollgraf, \textit{Fashion-MNIST: a novel image dataset for benchmarking machine learning algorithms}, arXiv:1708.07747, 2017.
\end{thebibliography}

\end{document}